Using Scagnostics to find Interesting Projections of Multivariate Data

### 2018-02-27 scagnostics.Rmd
Started adding Geozoo as example.

@@

@@

```{r load}
library(tourr)
library(tidyverse)
library(reshape2)
library(binostics) #https://github.com/uschiLaa/binostics
library(gridExtra)
library(wesanderson) #color palette
library(tictoc) #timer
library(mbgraphic) #Katrins package
library(dfoptim)
```

```{r tourpath}
data(mtcars)
t1 <- save_history(mtcars, grand_tour(2), max=4)
t1full <- as.list(interpolate(t1))
n <- length(t1full)
sc <- NULL
for (i in 1:n) {
  dprj <- as.matrix(mtcars) %*% t1full[[i]]
  sc <- rbind(sc, scagnostics(dprj))
}
sc <- as.data.frame(sc)
sc$t <- 1:n
sc <- sc %>% gather(index, value, -t)
```

```{r plotscag, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
ggplot(sc, aes(x=t, y=value)) + geom_line() + 
  facet_wrap(~index, ncol=1, scales = "free_y")
```

>>>>

```{r load}
library(tourr)
library(tidyverse)
library(reshape2)
library(binostics) #https://github.com/uschiLaa/binostics
library(gridExtra)
library(wesanderson) #color palette
library(tictoc) #timer
library(mbgraphic) #Katrins package
library(dfoptim)
library(GGally)
library(geozoo)
```

@@

## Dataset

As a test dataset we generate high dimensional geometrical shapes using the geozoo package (REF).
Concretely we consider five dimensional objects, and we combine a hollow shpere with radius 1 and a solid torus with radii (0.5, 0.5, 0.1,0.1). The resulting distributions can be seen in the scatter plot matrix below. In the following we combine both groups to study the resulting non-standard distribution.

```{r geozooDist, fig.width=6, fig.height=7, fig.cap="Scatter plot matrix of the five dimensional shapes, randomly selecting 100 points in each group. Group 1 (2) contains points selected on a hollow sphere (solid torus)."}
hSphere <- sphere.hollow(p=5,n=100)$points
sTorus <- torus(5,100, radius = c(0.5,0.5,0.1,0.1))$points
geoDF <- bind_rows(as.data.frame(hSphere),as.data.frame(sTorus),.id = "Group")
ggscatmat(geoDF, columns = 2:6, color="Group")
```

We first study the smoothness of the scagnostics measures calculated on a sequence of 2-d projections obtained via an interpolated grand tour path. The temporal trace of all nine measures is shown below, where each time step refers to an interpolation step with a target distance between interpolated bases is set to 0.05 radians. We observe sharp jumps in all nine measures, note however the large differences in overall scale between the measures. In particular "Clumpy", "Monotonic", "Sparse" and "Striated" show little sensitivity, as could be expected from the distributions in Figure (REF).

```{r tourpathGeozoo}
geoCombined <- select(geoDF,-Group)
t1 <- save_history(geoCombined, grand_tour(2), max=4)
t1full <- as.list(interpolate(t1))
n <- length(t1full)
sc <- NULL
for (i in 1:n) {
  dprj <- as.matrix(geoCombined) %*% t1full[[i]]
  sc <- rbind(sc, scagnostics(dprj))
}
sc <- as.data.frame(sc)
sc$t <- 1:n
scPairs <- sc
sc <- sc %>% gather(index, value, -t)
```

```{r plotscagGeozoo, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
ggplot(sc, aes(x=t, y=value)) + geom_line() + 
  facet_wrap(~index, ncol=1, scales = "free_y") +
  geom_vline(mapping = aes(xintercept=15, color="red")) +
  geom_vline(mapping = aes(xintercept=25, color="red")) +
  guides(color=FALSE)
```

```{r plotscagPairs, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections, as above. Scatter plot of scagnostics calculated on each projection. FIXME is this relevant? Maybe remove this plot."}
ggpairs(scPairs, columns = 1:9)
```

Two projections, t=15 and t=25, have been marked in Figure (can I add a reference?), as examples exhibiting jumps in several measures with respect to the previous and/or next projection. We compare the observed behaviour to the scatter plots obtained from the projected data in the Figure below.

ADD Figure next

@@
